Notes on Design and Implementation of Anglican

### 2015-07-16 paper/paper.tex
introduction drafted

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{minted}

@@

@@

\section{Runtime Library}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}

@@

Anglican is a probabilistic programming language integrated with
Clojure.  There are several ways to build a programming language
on top or besides another language.  The easiest to grasp is an
interpreter --- a program that reads a program, in its entirety
or line-by-line, and executes it by applying operational
semantics of a certain kind to the language. \textsc{Basic} is
famous for line-by-line interpreted implementations.

Another approach is to write a compiler, either to a virtual
architecture, so called p-code or byte-code, or to real
hardware. Here, the whole program is translated from the
`higher-level' source language to a `lower-level' object
language, which can be directly executed, either by hardware or
by an interpreter --- but the latter interpreter can be made
simpler and more efficient  than an interpreter for the source
language.

On top of these two approaches are methods in which a new
language is implemented `inside' another language of the same
level of abstraction. Different languages provide different
means for this; Lisp is notorious for the macro facility
that allows to extend the language almost without
restriction --- by writing \textit{macros}, one adds new
constructs to the existing language. There are several uses of
macros --- one is to extend the language \textit{syntax}, for
example, by adding new control structures; another is to keep
the existing syntax but alter the operational semantic --- the
way programs are executed and compute their outputs.

Anglican is implemented in just this way --- a macro facility
provided by Clojure, a Lisp dialect, is used both to add to
Clojure constructs that delimit probabilistic, rather than
deterministic, code, and to alter the operational semantics of
Clojure constructs inside probabilistic code fragments. Anglican
claims its right to count as a separate language because of the
ubiquitous probabilistic execution semantics rather than because
of a different syntax, which is actually an advantage rather
than a drawback --- Clojure programmers only need to know how to
specify the boundaries of Anglican programs, but can use
familiar Clojure syntax to write probabilistic code.

An implementation of Anglican must therefore address three issues:
\begin{itemize}
	\item the Clojure syntax to introduce probabilistic Anglican
		code inside Clojure modules;
	\item source-to-source transformation of Anglican programs
		into Clojure, so that probabilistic execution becomes
		possible;
	\item algorithms which run Clojure code, obtained by
		transforming Anglican programs, according to the
		probabilistic operational semantics.
\end{itemize}
The following sections explain the way Anglican is implemented,
from source-to-source transformation and syntactic wrappers to
inference algorithms which accept Clojure functions built from
anglican code as a parameter, and produce probabilistic outputs.

@@

\section{Definitions and Runtime Library}
